---
title:  'Homework 2'
subtitle: 'INFO 523'
author:
- name: Student -  [Your name]
  affiliation: [Your affiliation]
- name: Instructor -  Cristian Roman-Palacios
  affiliation: School of Information, University of Arizona, Tucson, AZ
tags: [R, RStudio, HW2]
output: html_document

---

---------------

### Objectives
This worksheet will assess your knowledge of basic commands in `R`. Please review the lectures, suggested readings, and additional resources **before** getting started on the HW as this document closely follows the provided materials.


---------------

## Additional files related to this HW
In `D2L`, you will find the following additional files associated to this homework:

- **1_Lab.RMD**: An annotated markdown style document to introduce `R`. If you’re not familiar with `R`, please **review** this document before going over this HW.

- **2_BasicR.R**: An annotated `R` script with different brief pactice questions (not graded).


#### Additional resources relevant to this HW
- **R Markdown**: Please review the basic R Markdown cheat sheet in case you have any questions regarding formatting the HW: https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.

- **R**: Please review the basic `R` cheat sheet in case you have any questions regarding the programming language: https://www.soa.org/globalassets/assets/Files/Edu/2018/exam-pa-base-r.pdf.

- **RStudio**: Additional cheat sheets written by RStudio to help with specific R packages: https://www.rstudio.com/resources/cheatsheets/

- **Datasets**: The following website has access to the relevant datasets from the recommended textbook: https://book.huihoo.com/introduction-to-statistical-learning/data.html


### Scores and grading policies

There is avery basic auto-grading system implemented at the end of the assignment. Use that workflow as a reference. However, please note that grades are **NOT exclusively based on your final answers**. I will be grading the overall structure and logic of your code. Feel free to use as many lines as you need to answer each of the questions. I also highly recommend and strongly encourage adding comments (`#`) to your code. Comments will certainly improve the reproducibility and readability of your submission. Commenting your code is also good coding practice. **Specifically for the course, you’ll get better feedback if the instructor is able to understand your code in detail.**


### Submission:
 This HW is due at the end of the same week that is released (**Sunday, 11:59 pm AZ time**). Please get in touch with the instructor if you’re (i) having issues opening the assignment, (2) not understanding the questions or (3) having issues submitting your assignment. Note that late submissions are subject to a penalty (see late work policies in the Syllabus). By the deadline, you should turn in a a `RMD` file (this file) **and** a rendered `HTML` (hint: knit your `rmd`; link: https://rmarkdown.rstudio.com/lesson-9.html). Answers to each question should be in the relevant block of code (see below). The instructor won't render your submission. **Re-name your file to `lastname_Hw1.RMD` before submitting to your repo**. Make sure that you can correctly render your submission without errors before turning anything in. If a given block of code is causing issues and you didn’t get to fix it, please use `r eval=FALSE `the in the relevant block and add comments. **This assignment must be submitted through our GitHub Classroom before the deadline.**


### Time commitment
Please reach out if you’re taking more than ~18h (~9h for the first week) to complete (1) this HW, (2) reading the book chapters, and (3) going over the lectures. I will be happy to provide accommodations if necessary. **Do not wait until the last minute to start working on this HW**. In most cases, working under pressure will certainly increase the time needed to answer each of these questions and the instructor might not be 100% available on Sundays to troubleshoot with you. Remember that you can sign up office hours with the instructor 3 times a week.


### Looking for help?
First, please go over the relevant readings for this week. Second, if you’re still struggling with any of the questions, do some independent research (e.g. stackoverflow is a wonderful resource). Don’t forget that your classmates will also be working on the same questions - reach out for help (check under the Discussion forum for folks looking to interact with other students in this class or start your own thread). Finally, the instructor will be happy to answer any questions during office hours. You can reach out to me by email (cromanpa94@arizona.edu) or simply schedule a 15 minute meeting through **Calendly** (https://calendly.com/cromanpa/15min). **Do not forget that the instructor holds office hours 3 times a week!!**


---------------
# Questions
Each question is worth a single point. Please answer only the questions labeled with the section that you’re enrolled in. 

#### Question 1

Consider a vector:

```{r}
x <- c(4,6,5,7,10,9,4,15)
```
Create a logical vector indicating which elements are `<7`.

```{r}
# BEGIN SOLUTION
q1 <- NULL #Your final answer
# END SOLUTION
```

#### Question 2


If:

```{r}
Age <- c(22, 25, 18, 20)
Name <- c("James", "Mathew", "Olivia", "Stella")
Gender <- c("M", "M", "F", "F")
```

, what is the R-code for getting the following output?

|   | Age | Name   | Gender |
|---|-----|--------|--------|
| 1 | 22  | James  | M      |
| 2 | 25  | Mathew | F      |


```{r}
# BEGIN SOLUTION
q2 <- NULL #Your final answer
# END SOLUTION
```


#### Question 3

Consider the vector:

```{r}
 x <- c(34, 56, 55, 87, NA, 4, 77, NA, 21, NA, 39)
```
How would you count the number of `NA` values in x?

a. `count(is.na(x))`
b. `length(is.na(x))`
c. `sum(is.na(x))`
d. `count(!is.na(x))`
e. `sum(!is.na(x))`

```{r}
# BEGIN SOLUTION
q3 <- NULL #Your final answer
# END SOLUTION
```


#### Question 4

If `x <- c("ss", "aa", "ff", "kk", "bb")`, what will be the third value in the index vector operation `x[c(2, 4, 4)]`?


```{r}
# BEGIN SOLUTION
q4 <- NULL #Your final answer
# END SOLUTION
```



#### Question 5

What will be the output for the R expression `(c(FALSE, TRUE)) || (c(TRUE, TRUE))`? Explain.

```{r}
# BEGIN SOLUTION
q5 <- NULL #Your final answer
# END SOLUTION
```

#### Question 6

Write an R expression that will return the positions of 3 and 7 in the vector `x <- c(1, 3, 6, 7, 3, 7, 8, 9, 3, 7, 2)`


```{r}
# BEGIN SOLUTION
q6 <- NULL #Your final answer
# END SOLUTION
```

#### Question 7

If `a <- list ("x"=5, "y"=10, "z"=15)`, how would you estimate the sum of all elements in a?


```{r}
# BEGIN SOLUTION
q7 <- NULL #Your final answer
# END SOLUTION
```

#### Question 8

Let `x <- list(a=5:10, c="Hello", d="AA")`, write an R statement to add a new item z = "NewItem" to the list `x`.

```{r}
# BEGIN SOLUTION
q8 <- NULL #Your final answer
# END SOLUTION
```


#### Question 9

Create a function to convert inches to centimeters. Assume the input data are in inches and you want to return the same data converted to cm. Your function must be called `q9.function`. Use `q9` to test your function with `3201 in`.

```{r}
# BEGIN SOLUTION
q9.function <- NULL #Your function answer
q9 <- NULL #Your final answer
# END SOLUTION
```

#### Question 10

The following vector represents the number of vacation days possessed by various individuals

```{r}
vd <- c(5, 7, 20, 1, 0, 0, 12, 4, 2, 2, 2, 4, 5, 6, 7, 10, 4)
```

Answer the following questions by saving your reponse in a named list with 3 elements:

- How many individuals are represented in the vector? (name in list=`totalInd`)
- How many vacation days total? (name in list=`totalVd`)
- How many individuals have four or less vacation days left? (name in list=`less4Vd`)

```{r}
q10 <- list(NA, NA, NA)
```


#### Question 11

The points below relate to the GitHub component reviewed this week. Please follow the instructions in order and provide the relevant links:

1. Clone the following repo (locally): https://github.com/rstudio/Intro
2. Create a new public repo (with any name) in your GitHub account. Push the same contents as in the original repo to your new repo. List the link to the repo in `q11.2`.

```{r}
# BEGIN SOLUTION
q11.2 <- NULL ##paste your link here (as a string)
# END SOLUTION
```

3. Create a new organization associated with your account (e.g. firstname_DM523). Below (`q11.3`), list the link to the organization.

```{r}
# BEGIN SOLUTION
q11.3 <- NULL ##paste your link here (as a string)
# END SOLUTION
```


4. Use the new organization to fork the repo you created in #2. List the link to the forked repo in `q11.4`.

```{r}
# BEGIN SOLUTION
q11.4 <- NULL ##paste your link here (as a string)
# END SOLUTION
```

5. From the forked repo (#4), select one of the scripts under Code/, add some annotations, commit, and push. Your forked should be >0 commit ahead of your otiginal repo in #2. List the link to the new commit in `q11.5`.


```{r}
# BEGIN SOLUTION
q11.5 <- NULL ##paste your link here (as a string)
# END SOLUTION
```

6. Again from the forked repo in your organization (#4), open a pull request to your original repo (#2). List the link to the pull request in `q11.6`.


```{r}
# BEGIN SOLUTION
q11.6 <- NULL ##paste your link here (as a string)
# END SOLUTION
```

## Some quick feedback

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
library(digest)
library(here)
load(here("tests", "ref_WS.RData"))
.grade(submission=list(q1, q2, q3, q4, q5, 
                       q6, q7, q8, q9, q10,
                       q11.2,q11.3,q11.4,q11.5,q11.6
                       )
       )
```


