---
title:  'Homework 2'
subtitle: 'INFO 523'
author:
- name: Student -  [Your name]
  affiliation: [Your affiliation]
- name: Instructor -  Cristian Roman-Palacios
  affiliation: School of Information, University of Arizona, Tucson, AZ
tags: [R, RStudio, HW2]
output: html_document
---

---------------

### Objectives
This worksheet will assess your knowledge of basic commands in `R`. Please review the lectures, suggested readings, and additional resources **before** getting started on the HW as this document closely follows the provided materials.


---------------

## Additional files related to this HW
In `D2L`, you will find the following additional files associated to this homework:

- **1_Lab.RMD**: An annotated markdown style document to introduce `R`. If you’re not familiar with `R`, please **review** this document before going over this HW. This document is related to the intro to `R` presented in the main textbook for this course.


- **2_BasicR.R**: An annotated `R` script with different questions (used in the extra credit for HW1). There is also a quick recording posted in `D2L` of the instructor reviewing the code in this script.


#### Additional resources relevant to this HW
- **R Markdown**: Please review the basic R Markdown cheat sheet in case you have any questions regarding formatting the HW: https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.

- **R**: Please review the basic `R` cheat sheet in case you have any questions regarding the programming language: https://www.soa.org/globalassets/assets/Files/Edu/2018/exam-pa-base-r.pdf.

- **RStudio**: Additional cheat sheets written by RStudio to help with specific R packages: https://www.rstudio.com/resources/cheatsheets/

- **Datasets**: The following website has access to the relevant datasets from the recommended textbook: https://book.huihoo.com/introduction-to-statistical-learning/data.html


#### The *Tidyverse*

I encourage students to check out functions from packages included in the `tidyverse` (https://www.tidyverse.org/) which greatly facilitates the productivity of novice coders. However, all the instruction will be delivered using `Base` `R`. The main textbook also uses base `R` only. I will be happy to grade your code regardless whether it uses base `R` or functions in the `tidyverse`. For some steps, other packages (such as `data.table`) are an appropriate alternative. Most if not all the questions in this HW can be answered using the `tidyverse`. Please check out the “accompanying” book to our main textbook that uses packages from the `tidyverse` instead of base (https://emilhvitfeldt.github.io/ISLR-tidymodels-labs/index.html).



### Scores

Please answer the questions from the section that you’re enrolled in (labeled as either **421** or **521**). Below is a summary of the total scores associated with this HW for undergraduates and graduate students.


- **ISTA421**: 9 points (undergraduate)
- **INFO521**: 12 points (graduate)
- **Extra credit** (ISTA421/INFO521): +3 points 


### Submission:
Please follow the instructions outlined below to submit your assignment. This HW is due at the end of the same week that is released (**Sunday, 11:59 pm AZ time**). Please get in touch with the instructor if you’re (i) having issues opening the assignment, (2) not understanding the questions or (3) having issues submitting your assignment. Note that late submissions are subject to a penalty (see late work policies in the Syllabus).


-**Homework 1**: Please turn in a single `RMD` file (this file) and a rendered `HTML` (hint: knit your `rmd`; link: https://rmarkdown.rstudio.com/lesson-9.html). Answers to each question should be in the relevant block of code (see below). Re-name your file to `lastname_Hw1.RMD` before submitting. Make sure that you can correctly render  your submission without errors before turning anything in. If a given block of code is causing issues and you didn’t get to fix it, please use `r eval=FALSE `the in the relevant block and add comments.

-**Extra credit**: An annotated R script. Re-name your file to `lastname_Hw1_EC.RMD` before submitting.

### Time commitment
Please reach out if you’re taking more than ~18h (~9h for the first week) to complete (1) this HW, (2) reading the book chapters, and (3) going over the lectures. I will be happy to provide accommodations if necessary. **Do not wait until the last minute to start working on this HW**. In most cases, working under pressure will certainly increase the time needed to answer each of these questions and the instructor might not be 100% available on Sundays to troubleshoot with you. Remember that you can sign up office hours with the instructor 3 times a week.



### Looking for help?
First, please go over the relevant readings for this week. Second, if you’re still struggling with any of the questions, do some independent research (e.g. stackoverflow is a wonderful resource). Don’t forget that your classmates will also be working on the same questions - reach out for help (check under the Discussion forum for folks looking to interact with other students in this class or start your own thread). Finally, the instructor will be happy to answer any questions during office hours. You can reach out to me by email (cromanpa94@arizona.edu) or simply schedule a 15 minute meeting through **Calendly** (https://calendly.com/cromanpa/15min). **Do not forget that the instructor holds office hours 3 times a week!!**

### Grading
Please note that grades are **NOT exclusively based on your final answers**. I will be grading the overall structure and logic of your code. Feel free to use as many lines as you need to answer each of the questions. I also highly recommend and strongly encourage adding comments (`#`) to your code. Comments will certainly improve the reproducibility and readability of your submission. Commenting your code is also good coding practice. **Specifically for the course, you’ll get better feedback if the instructor is able to understand your code in detail.**


---------------

# Questions
Each question is worth a single point. Please answer only the questions labeled with the section that you’re enrolled in. 

**NOTE**: If you’re enrolled in ISTA 421 and decide to answer questions labeled as 521, the instructor will grade them for extra credit. However, students enrolled in ISTA421 are **NOT** required to answer questions labeled as 521.


#### Question 1

Consider a vector:

```{r}
x <- c(4,6,5,7,10,9,4,15)
```

What is the value of: `c(4,6,5,7,10,9,4,15) < 7`


a. TRUE, FALSE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE
b. TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE
c. FALSE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE
d. TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, FALSE
e. TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE


```{r}
# BEGIN SOLUTION
q1 <- NULL #Your final answer
# END SOLUTION
```

#### Question 2


If:

```{r}
 Age <- c(22, 25, 18, 20)
 Name <- c("James", "Mathew", "Olivia", "Stella")
 Gender <- c("M", "M", "F", "F")
```

 then what is the R-code for getting the following output?

|   | Age | Name   | Gender |
|---|-----|--------|--------|
| 1 | 22  | James  | M      |
| 2 | 25  | Mathew | F      |


a.
DataFrame = data.frame(c(Age), c(Name), c(Gender))
subset(DataFrame, Gender == "M")
 
b.
DataFrame = data.frame(c(Age),c(Name),c(Gender))
subset(Gender=="M"), eval=FALSE 

c.
DataFrame = data.frame(Age,Name,Gender)
subset(DataFrame,Gender=="M")
 
d. 
DataFrame = data.frame(c(Age,Name,Gender))
subset(DataFrame,Gender=="M")


```{r}
# BEGIN SOLUTION
q2 <- NULL #Your final answer
# END SOLUTION
```


#### Question 3

Consider the vector:

```{r}
 x <- c(34, 56, 55, 87, NA, 4, 77, NA, 21, NA, 39)
```
How would you count the number of `NA` values in x?

a. `count(is.na(x))`
b. `length(is.na(x))`
c. `sum(is.na(x))`
d. `count(!is.na(x))`
e. `sum(!is.na(x))`

```{r}
# BEGIN SOLUTION
q3 <- NULL #Your final answer
# END SOLUTION
```


#### Question 4

If `x <- c("ss", "aa", "ff", "kk", "bb")`, what will be the third value in the index vector operation `x[c(2, 4, 4)]`?


```{r}
# BEGIN SOLUTION
q4 <- NULL #Your final answer
# END SOLUTION
```



#### Question 5

What will be the output for the R expression `(c(FALSE, TRUE)) || (c(TRUE, TRUE))`? Explain.

```{r}
# BEGIN SOLUTION
q5 <- NULL #Your final answer
# END SOLUTION
```

#### Question 6

Write an R expression that will return the positions of 3 and 7 in the vector `x <- c(1, 3, 6, 7, 3, 7, 8, 9, 3, 7, 2)`


```{r}
# BEGIN SOLUTION
q6 <- NULL #Your final answer
# END SOLUTION
```

#### Question 7

If `a <- list ("x"=5, "y"=10, "z"=15)`, how would you estimate the sum of all elements in a?


```{r}
# BEGIN SOLUTION
q7 <- NULL #Your final answer
# END SOLUTION
```

#### Question 8

Let `x <- list(a=5:10, c="Hello", d="AA")`, write an R statement to add a new item z = "NewItem" to the list `x`.

```{r}
# BEGIN SOLUTION
q7 <- NULL #Your final answer
q8 <- NULL
q9 <- NULL
q10 <- NULL
# END SOLUTION
```


#### Question 9



Now, complete the return statement below to create a function to convert inches to centimeters. Assume the input data are in inches and you want to return the same data converted to cm.


#### Question 10

Write a function to compute the standard deviation, use the four steps outlined above. Do not use the internal sd() function to check your work because it uses \( \sqrt{\frac{1}{N} \sum_{i=1}^N (x_i - \overline{x})^2} \), note the N-1 in the denominator. Your function is correct if you find a standard deviation of around 3.39.


#### Question 11

Create an object called agge that contains your age in years. Then reassign the object to a new object called age (e.g., correct the typo). Then remove the previous object from your environment and then calculate your age in days

#create agge object

#reassign agge value to age

#remove agge object

#Use math to calculate your age in days



#### Question 12

The following vector represents the number of vacation days possessed by various employees.

How many employees are represented in the vector?

How many vacation days total?

vacation_days <- c(5, 7, 20, 1, 0, 0, 12, 4, 2, 2, 2, 4, 5, 6, 7, 10, 4)

#how many employees are represented in the vector?

#how many vacation days total?


## Some quick feedback

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
library(digest)
library(here)
load(here("tests", "ref_WS.RData"))
.grade(submission=list(q1, q2, q3, q4, q5, 
                       q6, q7, q8, q9, q10
                       )
       )
```


